\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\RequirePackage{natbib}
\usepackage{mathrsfs}
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}


\usepackage{ulem}
\usepackage{setspace}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{prop}{Proposition}

\newcommand{\m}{\mathcal}

\title{Paper Summary}
\author{Song Wang\\
Statistics, UW-Madison}
%\date{}                                           % Activate to display a given date or no date


\begin{document}
\maketitle
{\bf Physician Patient-Sharing Networks and the Cost and Intensity of Care in US Hospital}

-- Trying to assess how the structure of patient-sharing networks of physician within hospitals might contribute to variation in the cost and intensity.

-- data: patient-shareing network from 2006 medicare data + information about hospitals and physisian from AMA 2006 annual  survey.

-- Method: Assigning physisian to primray hospital; for each hospital, Use measurements ( adjusted degree, between centrality) of the network structure, and some other hospital control variables as predictors; hospital measures of spending and care intensity in the past two years as response variable. Run regression.

{\bf  Physician Networks and Ambulatory Care-sensitive Admissions}
-- Objective: Determin the relationship between  rate ambulatory care-sensitive hospital admissions varies across physician networks /network characteristics..

-- Data : 2008 medicare , 987,000 beneficiaries in 5 states.

--  Confine instead of looking the patient-sharing network spanned by physicians across the country, look at small networks among physicians within small communities -- PPC physician practice community. PPCs are found through network algorithm.   

    For each PPCs, it has certain beneficiaries Beneficiary Annual Summary File), certain physicians (AMA masterfile). + phsician adjusted degree and between centrality. some PPCs level characteristics are calculated from physicians ( percentage of primary care physicians, mean adjusted degree etc)
    
    regression on the beneficiary level, with its own characteristics and community characteristics. Use the ACSA's as response ( $\#$ of Ambulatory Care-sensitive Admissions).









\end{document}  






