#### Study the asymmetry of the edges in the specialty referral matrix. 

We are studying a small graph, where each node is a primary specialty (e.g. PREVENTATIVE MEDICINE).  Then, an "edge" from one specialty to another specialty represents a referral from a doctor of the first type, to a doctor of the second type.  In a sample of 100k edges, there were 74 unique primary specialties. To accumulate edge strength, I used "two logs" as was done in the zip code graph.  You can [see my code here](code/smallSpecialtyReferralGraph.R).  You can download the data from the following url.
```{r}
rm(list = ls())
library(Matrix)
library(igraph)
load(url("http://pages.stat.wisc.edu/~karlrohe/netsci/data/specialty.RData"))
A = as.matrix(A) # it used to be a sparse matrix "dgCMatrix", even though it isn't that sparse. 
rs = rowSums(A); cs = colSums(A)
plot(rs,cs, xlab= "out degree", ylab = "in degree", col = "white")
text(rs, cs, labels=colnames(A), cex = .5)
ors = order(rs)
cbind(colnames(A)[ors], round(rs[ors],1))
```


Are edges reciprocal? 
```{r}
plot(A,t(A), main = "non-reciprocated edges at low degrees.")

bA = matrix(0, nrow = nrow(A), ncol = ncol(A))
bA[A>.5] = 1
nonrecA = bA*(1-t(bA))
nonrecEdges = which(nonrecA == 1, arr.ind = T)
# the first column makes some referrals to the second column, but it is much less likely in the other direction.
cbind(colnames(A)[nonrecEdges[,1]], colnames(A)[nonrecEdges[,2]])
```

What is the page rank? Imagine a Markov chain on the specialties.  You get randomly referred around this network.  Where are you likely to land?
```{r}
g= graph.adjacency(A,mode= "directed", weighted = T)
pr = page.rank(g,directed = T)
sort(pr$vec)
```

Before reading the next steps, you might read the [lecture notes on the eigendecomposition and the svd](eigenSVDlecture.pdf).  
The algorithm "di-sim" explored here is described in [this paper](http://arxiv.org/abs/1204.2296). 

```{r}
Drow = Diagonal(n = nrow(A), 1/sqrt(rs + mean(rs)))
Dcol = Diagonal(n = ncol(A), 1/sqrt(cs + mean(cs)))
L = Drow%*%A%*%Dcol
s = svd(L)
plot(s$d[-1])
```

```{r}
k = 3
u = s$u[,1:k]; v = s$v[,1:k]
n = nrow(A)
un = t(apply(u,1, function(x) return(x/sqrt(sum(x^2)+1/n))))
vn = t(apply(v,1, function(x) return(x/sqrt(sum(x^2)+1/n))))
km = kmeans(rbind(un,vn), centers = k,nstart = 100)
uclust = km$clust[1:n]
vclust = km$clust[-(1:n)]
table(uclust,vclust)
```
If this network was highly symmetric (i.e. reciprocal) what would that table look like?
```{r}
nm = colnames(A)
nm[uclust == 1]
nm[vclust == 1]
```
Note that $u$ corresponds to sending patterns and $v$ corresponds to receiving patterns.